|

### 2020-01-09 inst/manuscript/candidaev_supplement.Rmd
Add supplementary figures S4 S5
Figure S4A is a venn diagram comparing yeast EV proteins with biofilm
EV proteins.

Figure S4B is a volcano plot showing enrichment of proteins common to
yeast and biofilm EVs.

Data underlying figure S4 are given in supplementary data S6.

Figure S5A is a GO enrichment analysis performed on yeast EV proteins
that are exclusive or enriched compared to biofilm EVs.

Figure S5B is  a GO enrichment analysis performed on biofilm EV
proteins that are exclusive or enriched compared to yeast EVs.

Data underlying figure S5 are given in supplmentary data S7

>>>>

```{r figureS4Data}
#### DAY286 yeast EV proteins ####
# filter for proteins quantified in min 2/3 reps of yeast EV or WCL
# then convert to a data.frame and extract protein acessions from row names
yeast_filt3 <- filter_na2(yeast_lfq2, logic = "or", op = "<=",
                          pat1 = "EV", val1 = 1,
                          pat2 = "W", val2 = 1) %>%
  as.data.frame() %>%
  mutate(id = stringr::str_extract(rownames(.), "[^;]+")) %>%
  select(id, everything())

#### DAY286 biofilm EV proteins ####
# filter for proteins quantified in min 4/5 reps of biofilm EV or WCL
# then convert to a data.frame and extract protein acessions from row names
biofilm_filt2 <- filter_na2(biofilm_lfq, logic = "or", op = "<=",
                           pat1 = "EV", val1 = 1,
                           pat2 = "W", val2 = 1) %>%
  as.data.frame() %>%
  mutate(id = stringr::str_extract(rownames(.), "[^;]+")) %>%
  select(id, everything())

#### Combine yeast EV and biofilm EV data ####
# Combine the yeast and biofilm datasets by matching Uniprot accessions
# Remove the WCL columns
# Filter for proteins with at least 1 valid LFQ value
# Convert to numeric matrix
ev_comb <- full_join(yeast_filt3, biofilm_filt2, by = "id") %>%
  select(id, contains("EV")) %>%
  tibble::column_to_rownames(var = "id") %>%
  filter_na(op = "<=", pat = "EV", val = 7) %>%
  as.matrix()

# Rename the columns to something useful
colnames(ev_comb) <- c("Y_EV1", "Y_EV2", "Y_EV3",
                       "B_EV1", "B_EV2", "B_EV3", "B_EV4", "B_EV5")

# Check the yeast EV and biofilm EV proteins match the counts in Figure 2A
b_ev2 <- filter_na(ev_comb, op = "<=", pat = "B_EV", val = 4)
y_ev2 <- filter_na(ev_comb, op = "<=", pat = "Y_EV", val = 2)

# check_venn <- plot_venn(list("Y_EV" = y_ev, "B_EV" = b_ev),
#                         use_uniprot = TRUE, type = "plot",
#                         fontfamily = "sans",
#                         cex = 0.8,
#                         cat.fontfamily = "sans",
#                         cat.just = list(c(0.5,1.7), c(0.5,1.5)),
#                         margin = 0.05)

# Import the yeast EV vs biofilm EV experimental design required for limma
ev_exp <- read.table("supplementary_data/Y_EV_vs_B_EV_expdesign.txt",
                     sep = "\t", header = TRUE, stringsAsFactors = FALSE)

#### Data processing ####
# Normalise log2 LFQ data
ev_norm <- normalizeCyclicLoess(ev_comb)

# Won't impute with so many missing values
# Separate EV/WCL exclusive proteins (don't impute) from
# EV/WCL common proteins (impute)
ev_excl <- filter_na2(ev_norm, logic = "or", op = "==",
                      pat1 = "Y_EV", val1 = 3,
                      pat2 = "B_EV", val2 = 5)

ev_both <- filter_na2(ev_norm, logic = "and", op = "<=",
                      pat1 = "Y_EV", val1 = 2,
                      pat2 = "B_EV", val2 = 4)

# Use bpca method from pcaMethods for missing value imputation
ev_imp <- pcaMethods::pca(ev_both, method = "bpca")
ev_imp <- pcaMethods::completeObs(ev_imp)

ev_de <- rbind(ev_excl, ev_imp)

# see limma user guide section 9.2 for more info about DE
# create design matrix
ev_samp <- data.frame(T = c(rep("Y_EV", 3), rep("B_EV", 5)))

ev_design <- stats::model.matrix(~ 0 + T, data = ev_samp)
colnames(ev_design) <- c("B_EV", "Y_EV")

# define sample comparisons of interest
ev_contrasts <- c("Y_EV - B_EV")

# make all pair-wise comparisons between EV and WCL
# and perform limma::eBayes()
ev_efit <- limma_eBayes(ev_de, design = ev_design, contrasts = ev_contrasts)

# extract DE results
# adj.p.val cutoff is 0.01, log2FC cutoff is 1
ev_res <- get_results(efit = ev_efit,
                      mat = ev_de,
                      p_val = 0.01,
                      lfc = 1,
                      type = "individual")[[1]]

## Construct figure S4 components
# Fig S4A = compare DAY286 yeast EV and biofilm EV proteins via venn diagram
fS4_a <- plot_venn(list("DAY286\nYeast EV" = y_ev2, 
                             "DAY286\nBiofilm EV" = b_ev2),
                        use_uniprot = TRUE, type = "plot",
                        fontfamily = "sans",
                        cex = 0.8,
                        cat.fontfamily = "sans",
                        cat.just = list(c(0.5,1.7), c(0.5,1.5)),
                        margin = 0.05) %>% 
  as_ggplot() + 
  theme(plot.margin = unit(c(0,1.5,0,1.5), "cm"))

# Fig S4B = compare compare enrichment of common proteins to DAYY286 yeast EVs
# and DAY286 biofilm EVs via volcano plot
fS4_b <- plot_volcano(data = ev_res, 
                        p_val = adj.P.Val, 
                        log2fc = logFC,
                        group = group, 
                        use_labels = TRUE,
                        label_fun = ifelse(logFC >= 5 | logFC <= -5, 
                                           CGD_gene_name, ""), 
                        point_size = 0.6, 
                        label_size = 2, 
                        x_lim = c(-8, 8), 
                        y_lim = c(0, 8), 
                        legend_params = c("left", 11, 8), 
                        axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("Yeast EV enriched", 
                                 "Biofilm EV enriched", 
                                 "Not significant"), 
                      values = c("#0072B2", 
                                 "#999999", 
                                 "#D55E00"), 
                      breaks = c("Y_EV up", "B_EV up", "not sig"), 
                      name = element_blank()) + 
  annotate("text", x = 5.5, y = 0.5, label = "bold('DAY286')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('136')", 
           parse = TRUE, colour = "#D55E00") + 
  annotate("text", x = -3.5, y = 7, label = "bold('129')", parse = TRUE, 
           colour = "#0072B2") + 
  annotate("text", x = 0, y = 6, label = "bold('168')", parse = TRUE, 
           colour = "#999999") + 
  theme(legend.position = "right")

# arguments for figure S4 export
fS4_args <- list(tiff = list(type = "cairo", compression = "lzw"), 
                postscript = list(bg = "white", fonts = "sans"))
```

```{r figureS4, fig.cap=fS4_cap, fig.height=6, fig.width=5, fig.path='figures/', dev=c('pdf', 'postscript', 'tiff'), dev.args=fS4_args, dpi=600}
fS4_cap <- c("\\textbf{MAIN} Sub.")

# draw figure S4
ggarrange(fS4_a, fS4_b, 
          labels = c("A", "B"), 
          ncol = 1, 
          nrow = 2)
```

```{r figureS5Data}
#### Fig S5A = DAY286 yeast EV enr and ex proteins GO enrichment ####
# Copy protein names for pasting into FungiFun2 online GO enrichment tool
# FungiFun2 parameters:
# 1. Strain = Candida albicans (strain SC5314/ATCC MYA-2876)
# 2. Classification ontology = GO
# 3. Input IDs = paste protein names here
# 4. Advanced options = default except significance level = 0.01

# Yeast EV exclusive proteins
# Write to clipboard, paste into FungiFun2, and submit
yev_ex <- ev_res %>%
  filter(group == "Y_EV ex") %>%
  pull(CGD_gene_name)

# Yeast EV enriched proteins
# Write to clipboard, paste into FungiFun2, and submit
yev_up <- ev_res %>%
  filter(group == "Y_EV up") %>%
  pull(CGD_gene_name) %>%
  stringi::stri_replace_all_regex(.,
                                  c("MDH1"),
                                  c("CaO19.7481"),
                                  vectorize_all = FALSE)

# Import FungiFun2 functional (GO) enrichment results

# Yeast EV exclusive and enriched proteins
yev_go <- process_fungifun("supplementary_data/GO_supplementary/yev_ex_up.csv") %>%
  mutate(Prot.found = stringi::stri_replace_all_regex(.$Prot.found,
                                                      c("CaO19.7481"),
                                                      c("MDH1"),
                                                      vectorize_all = FALSE))

# define the colour palette
fS5_pal <- c("#6D2077", "#E4002B", "#FFB81C")

# define figure S5A
fs5_a <- plot_go_bar(yev_go, c(1:10, 17:26, 31:40), fS5_pal) + 
  facet_grid(rows = vars(GO.type), scales = "free") + 
  theme(legend.position = "none")

#### Fig S5B = DAY286 biofilm EV enr and ex proteins GO enrichment ####
# Copy protein names for pasting into FungiFun2 online GO enrichment tool
# FungiFun2 parameters:
# 1. Strain = Candida albicans (strain SC5314/ATCC MYA-2876)
# 2. Classification ontology = GO
# 3. Input IDs = paste protein names here
# 4. Advanced options = default except significance level = 0.01

# Biofilm EV exclusive proteins
# Write to clipboard, paste into FungiFun2, and submit
bev_ex <- ev_res %>%
  filter(group == "B_EV ex") %>%
  pull(CGD_gene_name) %>%
  stringi::stri_replace_all_regex(.,
                                  c("ALS3", "ARO2", "HOM3",
                                    "OSM1", "PLB3", "FBP1",
                                    "PST2", "STE23"),
                                  c("CaO19.1816", "CaO19.1986", "CaO19.1235",
                                    "CaO19.6882", "CaO19.6594", "CaO19.6178",
                                    "CaO19.3612", "CaO19.5561"),
                                  vectorize_all = FALSE)

# Biofilm EV enriched proteins
# Write to clipboard, paste into FungiFun2, and submit
bev_up <- ev_res %>%
  filter(group == "B_EV up") %>%
  pull(CGD_gene_name) %>%
  stringi::stri_replace_all_regex(.,
                                  c("PEP1"),
                                  c("CaO19.3767"),
                                  vectorize_all = FALSE)

# Import FungiFun2 functional (GO) enrichment results

# Biofilm EV exclusive and enriched proteins
bev_go <- process_fungifun("supplementary_data/GO_supplementary/bev_ex_up.csv") %>%
  mutate(Prot.found = stringi::stri_replace_all_regex(.$Prot.found,
                                                      c("CaO19.1816", "CaO19.1986", "CaO19.1235",
                                                        "CaO19.6882", "CaO19.6594", "CaO19.6178",
                                                        "CaO19.3612", "CaO19.5561", "CaO19.3767"),
                                                      c("ALS3", "ARO2", "HOM3",
                                                        "OSM1", "PLB3", "FBP1",
                                                        "PST2", "STE23", "PEP1"),
                                                      vectorize_all = FALSE))

# define figure S5B
fs5_b <- plot_go_bar(bev_go, c(1:10, 35:44, 54:63), fS5_pal) + 
  facet_grid(rows = vars(GO.type), scales = "free") + 
  theme(legend.position = "none")

# arguments for figure S5 export
fS5_args <- list(tiff = list(type = "cairo", compression = "lzw"), 
                postscript = list(bg = "white", fonts = "sans"))
```

```{r figureS5, fig.cap=fS5_cap, fig.height=7, fig.width=7, fig.path='figures/', dev=c('pdf', 'postscript', 'tiff'), dev.args=fS5_args, dpi=600}
fS5_cap <- c("\\textbf{MAIN} Sub.")

# draw figure S5
ggarrange(fs5_a, fs5_b, 
          labels = c("A\nYeast EV", 
                     "B\nBiofilm EV"), 
          ncol = 1, 
          nrow = 2, 
          align = "hv")
```
